Testing the Evolution of the Correlations between Supermassive Black Holes and their Host Galaxies using Eight Strongly Lensed Quasars

### 2020-08-14 manuscript/manuscript.tex
Merge with TT's edits

% Abstract of the paper
\begin{abstract}
One of the main challenges in using high redshift active galactic nuclei to study the correlations between the mass of the supermassive Black Hole (\mbh) and the properties of their active host galaxies is instrumental resolution. Strong lensing magnification effectively increases instrumental resolution and thus helps to address this challenge. In this work, we study eight strongly lensed active galactic nuclei (AGN) with deep {\it Hubble Space Telescope} (\hst) imaging, using the lens modelling code \lenstronomy\ to reconstruct the image of the source. Using the reconstructed brightness of the host galaxy, we infer the host galaxy stellar mass based on stellar population models.  \mbh\ are estimated from broad emission lines using standard methods. Our results are in good agreement with recent work based on non-lensed AGN, \blue{demonstrating the great potential of using strongly lensed AGNs to extend the study of the correlations to higher redshifts.}
At the moment, the sample size of lensed AGN is small and thus they provide mostly a consistency check on systematic errors related to resolution for the non-lensed AGN. However, the number of known lensed AGN is expected to increase dramatically in the next few years, through dedicated searches in ground and space based wide field surveys, and they may become a key diagnostic of black hole and galaxy co-evolution. 
\end{abstract}

@@

\blue{The inferred photometry of the host galaxy for all the eight systems are listed in Table~\ref{tab:host_measure}, (2)$-$(6) columns. The detailed information of the fitting process for each system is in Appendix~\ref{sec:photometry}.}

\begin{table*}
\renewcommand{\arraystretch}{1.25}
%\setlength{\tabcolsep}{20pt}
\centering
  \begin{threeparttable}
\caption{Summary of lensed AGN host inference.}\label{tab:host_measure}    
%\resizebox{12cm}{!}{
     \begin{tabular}{cccccccc}
\hline     
(1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) \\    
Object ID & intrinsic magnitude & magnitude & Host Flux Ratio (to Total) & \reff\ & \sersic\ $n$ & stellar population & $\log (M_{*}$)  \\
 & (source plane) & (image plane) & ( $\%$, Total = Host + AGN) & (arcsec) & & age (Gyr) & (M$_{\odot}$) \\ \hline
HE0435 & $21.49\substack{+0.40\\-0.29}$ & $18.58\substack{+0.30\\-0.23}$ & $36.0\pm11.1$ & $0.28\pm0.02$ & $2.71\pm0.20$ & $1.50$ & $10.91$ \\
RXJ1131$_{\rm bulge}$ & $21.80\substack{+0.23\\-0.19}$ & $18.70\substack{+0.07\\-0.06}$ & $7.1\pm1.4$ & $0.13\pm0.02$ & fix to 4 & $3.00$ & $10.39$ \\
RXJ1131$_{\rm disk}$ & $19.33\substack{+0.17\\-0.15}$ & $17.14\substack{+0.08\\-0.07}$ & $69.2\pm10.1$ & $0.90\pm0.06$ & fix to 1 & $1.50$ & $11.08$ \\
WFI2033 & $21.78\substack{+0.28\\-0.23}$ & $19.07\substack{+0.35\\-0.26}$ & $19.6\pm4.5$ & $0.28\pm0.02$ & $0.52\pm0.01$ & $0.625$ & $10.51$ \\
SDSS1206 & $21.31\substack{+0.23\\-0.19}$ & $18.30\substack{+0.05\\-0.05}$ & $33.3\pm6.4$ & $0.11\pm0.02$ & $4.57\pm0.53$ & $0.625$ & $10.77$ \\
HE1104 & $21.25\substack{+0.16\\-0.14}$ & $19.16\substack{+0.02\\-0.02}$ & $14.0\pm2.0$ & $0.27\pm0.02$ & $1.05\pm0.04$ & $0.625$ & $11.05$ \\
SDSS0246 & $23.44\substack{+0.28\\-0.22}$ & $20.85\substack{+0.08\\-0.07}$ & $4.0\pm0.9$ & $0.44\pm0.08$ & $4.96\pm0.08$ & $0.625$ & $10.75$ \\
HS2209 & $20.72\substack{+0.26\\-0.21}$ & $19.20\substack{+0.04\\-0.04}$ & $12.5\pm2.7$ & $1.96\pm1.28$ & $3.15\pm1.40$ & $1.00$ & $11.04$ \\
HE0047 & $22.92\substack{+0.48\\-0.33}$ & $20.37\substack{+0.20\\-0.17}$ & $2.3\pm0.8$ & $0.32\pm0.15$ & $4.18\pm0.75$ & $0.625$ & $10.91$ \\
\hline
\end{tabular}
%}
\begin{tablenotes}
      \small
      \item Note: $-$ Inference of the host galaxy properties. Column (2)-(6): photometry derived using the imaging data is listed Table~\ref{data_set}. Column (7): adopted age of stellar population with solar metallicity. Column (8): inferred stellar mass. \blue{The uncertainty level for \mstar\ is 0.2 dex}.
\end{tablenotes}    
\end{threeparttable}
\end{table*}

@@

{\bf HE0435} ~ {\color{blue}
%In addition to the F160W data, this system has been observed by \hst\ through filters F814W and F555W. The multi-band imaging data provides us with the opportunity to estimate the host color and improve the stellar mass estimate. We adopted the lensing modelling approach described in Section~\ref{sec:photometry} to infer the host photometry through the F814W and F555W bands. Based on the three band host magnitudes, we find that a 1.5~Gyr age and solar metallicity stellar population provides a good match to the color.
In addition to WFC3/F160W, HE0435 has been observed through bands ACS/F814W and ACS/F555W (GO-9744; PI: C. S. Kochanek). Our aim is to derive the brightness of the host galaxy through all the three bands to investigate the color and stellar population in the image plane. We use the modelling approach introduced in Section~\ref{sec:photometry} to perform this inference for the F814W and F555W bands. To save computer time, we only use the {\it PSF iteration} approach. We also fix the lens mass slope value to 1.9 since it is closer to the inference by~\citet[][i.e., $\gamma\sim1.93$]{Wong2017}. The inference of the fittings are shown in Figure~\ref{fig:app_HE0435}-(a), (b). Having obtained the magnitude of the lensed host in the three bands, we use the {\sc Gsf} package\footnote{\url{https://github.com/mtakahiro/gsf}}~\citep{Morishita2019} to perform the SED fitting. A set of age grids (up to 3.0 Gyr) are used in this fit with a constant SFR and a flexible form for star formation histories. Note that there is a degeneracy in the age and metallicity; however, this degeneracy has little effect on the \mstar\ inference~\citep{Bell2001}. In this work, we fix the metallicity to infer the age, and the uncertainty of the age is not considered. Finally, a stellar population with 1.5~Gyr of age and solar metallicity provides the best fit to the colors, as shown in Figure~\ref{fig:app_HE0435}-(c). This stellar population is used to estimate the host stellar mass.

@@

Considering the simulations by~\citet{Ding2017a} and the fact that we are able to obtain very consistent host magnitudes for the HE0435, RXJ1131 and WFI2033 using the independent approaches, the fidelity of the inferred magnitude is expected to be well characterized by the quoted uncertainties.
\blue{
Most of our systems do not have multi-band information to constrain the stellar template (e.g., star formation histories). We adopted a simple stellar population template to calculate the \mstar.
The uncertainties of the stellar populations could translate the substantial uncertainties into the \mstar. According to the investigation by \citet{Bell2001}, we estimate a typical level of 0.2~dex for the \mstar\ uncertainties.}

@@

\blue{
Note that this apparent evolution is obtained directly using the observed sample, which does not consider the selection effect yet~\citep{Tre++07,Schulze2011,Bennert++2011, Schulze2014,Park15}. For instance, adopting the framework introduced in~\citet{Schulze2015}, D20 estimated an expected bias of $+2.1$~dex in their observed \mbh\ for their 32 AGN sample, and the offset would be reduced to $\gamma=0.50\pm0.25$ after correcting for the selection bias. While in this work, the
lensed AGN systems were selected for time-delay cosmography work, based on the availability of a time delay and the known detectability of the host galaxy. This is a different selection function than that used by D20. It is thus reassuring that the two samples present the same apparent evolution. Naturally, inferring the true underlying evolution requires modelling the selection function (as studied by D20), which is not justified given the small sample size of lensed quasars sample. At this stage, the lensed quasars should thus be considered as a check on possible systematic measurement error in the D20 analysis rather than a stand-alone measurement.}

@@

We directly compare our sample to the recent measurements by~\citet[][D20]{Ding2020a}, who used the same approach to derive the \mstar\ and calibrated the \mbh\ using consistent recipes. The \mbh-\mstar\ correlation and its evolution with cosmic time {\bf are} in excellent agreement with the results obtained by D20, as shown in Figure~\ref{fig:scaling_relation}. \blue{Currently, the sample size of lensed AGN (i.e., eight systems) is quite limited; as a result, their selection effect has not been considered. Nevertheless, the present analysis of this good agreement with D20 in the \mbh-\mstar\ correlations strengthens conclusion in D20}, which can be summarized as follows. First, the growth of the supermassive black hole \blue{mildly} predates that of its host galaxy during their co-evolution, even when considering total stellar mass, as it is often done at high redshift. However, the actual morphologies of these hosts are likely to be more complex, including a disk and bulge component, as in the case of  RXJ1131. In contrast, the galaxies in the local sample are typically bulge-dominated and the bulge mass is adopted as their  \mstar. Thus, the reported evolution is weaker than what would be inferred by comparing \mbh\ to bulge \mstar\ at all redshifts \citep{Bennert++2011}. Taken together, these results are consistent with a scenario in which the stellar mass is transferred from disk to the bulge at a faster rate than the growth of \mbh\ since $z\sim2$.

>>>>

% Abstract of the paper
\begin{abstract}
One of the main challenges in using high redshift active galactic nuclei to study the correlations between the mass of the supermassive Black Hole (\mbh) and the properties of their active host galaxies is instrumental resolution. Strong lensing magnification effectively increases instrumental resolution and thus helps to address this challenge. In this work, we study eight strongly lensed active galactic nuclei (AGN) with deep {\it Hubble Space Telescope} (\hst) imaging, using the lens modelling code \lenstronomy\ to reconstruct the image of the source. Using the reconstructed brightness of the host galaxy, we infer the host galaxy stellar mass based on stellar population models.  \mbh\ are estimated from broad emission lines using standard methods. Our results are in good agreement with recent work based on non-lensed AGN, \blue{demonstrating the potential of using strongly lensed AGNs to extend the study of the correlations to higher redshifts.}
At the moment, the sample size of lensed AGN is small and thus they provide mostly a consistency check on systematic errors related to resolution for the non-lensed AGN. However, the number of known lensed AGN is expected to increase dramatically in the next few years, through dedicated searches in ground and space based wide field surveys, and they may become a key diagnostic of black hole and galaxy co-evolution. 
\end{abstract}

@@

\blue{The inferred photometric properties of the host galaxy for all the eight systems are listed in Table~\ref{tab:host_measure}, (2)$-$(6) columns. Detailed information on the fitting process for each system is given in Appendix~\ref{sec:photometry}.}

\begin{table*}
\renewcommand{\arraystretch}{1.25}
%\setlength{\tabcolsep}{20pt}
\centering
  \begin{threeparttable}
\caption{Summary of lensed AGN host inference.}\label{tab:host_measure}    
%\resizebox{12cm}{!}{
     \begin{tabular}{cccccccc}
\hline     
(1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) \\    
Object ID & intrinsic magnitude & magnitude & Host Flux Ratio (to Total) & \reff\ & \sersic\ $n$ & stellar population & $\log (M_{*}$)  \\
 & (source plane) & (image plane) & ( $\%$, Total = Host + AGN) & (arcsec) & & age (Gyr) & (M$_{\odot}$) \\ \hline
HE0435 & $21.49\substack{+0.40\\-0.29}$ & $18.58\substack{+0.30\\-0.23}$ & $36.0\pm11.1$ & $0.28\pm0.02$ & $2.71\pm0.20$ & $1.50$ & $10.91$ \\
RXJ1131$_{\rm bulge}$ & $21.80\substack{+0.23\\-0.19}$ & $18.70\substack{+0.07\\-0.06}$ & $7.1\pm1.4$ & $0.13\pm0.02$ & fix to 4 & $3.00$ & $10.39$ \\
RXJ1131$_{\rm disk}$ & $19.33\substack{+0.17\\-0.15}$ & $17.14\substack{+0.08\\-0.07}$ & $69.2\pm10.1$ & $0.90\pm0.06$ & fix to 1 & $1.50$ & $11.08$ \\
WFI2033 & $21.78\substack{+0.28\\-0.23}$ & $19.07\substack{+0.35\\-0.26}$ & $19.6\pm4.5$ & $0.28\pm0.02$ & $0.52\pm0.01$ & $0.625$ & $10.51$ \\
SDSS1206 & $21.31\substack{+0.23\\-0.19}$ & $18.30\substack{+0.05\\-0.05}$ & $33.3\pm6.4$ & $0.11\pm0.02$ & $4.57\pm0.53$ & $0.625$ & $10.77$ \\
HE1104 & $21.25\substack{+0.16\\-0.14}$ & $19.16\substack{+0.02\\-0.02}$ & $14.0\pm2.0$ & $0.27\pm0.02$ & $1.05\pm0.04$ & $0.625$ & $11.05$ \\
SDSS0246 & $23.44\substack{+0.28\\-0.22}$ & $20.85\substack{+0.08\\-0.07}$ & $4.0\pm0.9$ & $0.44\pm0.08$ & $4.96\pm0.08$ & $0.625$ & $10.75$ \\
HS2209 & $20.72\substack{+0.26\\-0.21}$ & $19.20\substack{+0.04\\-0.04}$ & $12.5\pm2.7$ & $1.96\pm1.28$ & $3.15\pm1.40$ & $1.00$ & $11.04$ \\
HE0047 & $22.92\substack{+0.48\\-0.33}$ & $20.37\substack{+0.20\\-0.17}$ & $2.3\pm0.8$ & $0.32\pm0.15$ & $4.18\pm0.75$ & $0.625$ & $10.91$ \\
\hline
\end{tabular}
%}
\begin{tablenotes}
      \small
      \item Note: $-$ Inference of the host galaxy properties. Column (2)-(6): photometry derived using the imaging data is listed Table~\ref{data_set}. Column (7): adopted age of stellar population with solar metallicity. Column (8): inferred stellar mass. \blue{The uncertainty of \mstar\ is estimated to be 0.2 dex}.
\end{tablenotes}    
\end{threeparttable}
\end{table*}

@@

{\bf HE0435} ~ {\color{blue}
%In addition to the F160W data, this system has been observed by \hst\ through filters F814W and F555W. The multi-band imaging data provides us with the opportunity to estimate the host color and improve the stellar mass estimate. We adopted the lensing modelling approach described in Section~\ref{sec:photometry} to infer the host photometry through the F814W and F555W bands. Based on the three band host magnitudes, we find that a 1.5~Gyr age and solar metallicity stellar population provides a good match to the color.
In addition to WFC3/F160W, HE0435 has been observed through bands ACS/F814W and ACS/F555W (GO-9744; PI: C. S. Kochanek). Our aim is to derive the brightness of the host galaxy through all the three bands to investigate the color and stellar population in the image plane. We use the modelling approach introduced in Section~\ref{sec:photometry} to perform this inference for the F814W and F555W bands. To save computer time, we only use the {\it PSF iteration} approach. We also fix the lens mass slope value to 1.9 since it is closer to the inference by~\citet[][i.e., $\gamma\sim1.93$]{Wong2017}. The inference of the fittings are shown in Figure~\ref{fig:app_HE0435}-(a), (b). Having obtained the magnitude of the lensed host in the three bands, we use the {\sc Gsf} package\footnote{\url{https://github.com/mtakahiro/gsf}}~\citep{Morishita2019} to perform the SED fitting. A range of ages (up to 3.0 Gyr) is used in this fit with a constant SFR and a flexible form for star formation histories. Note that there is a well-known degeneracy between age and metallicity; however, this degeneracy has little effect on the \mstar\ inference~\citep{Bell2001}. In this work, we fix the metallicity to infer the age. In the end, a stellar population with 1.5~Gyr of age and solar metallicity provides the best fit to the colors, as shown in Figure~\ref{fig:app_HE0435}-(c), although this choice is by no means unique. This stellar population is used to estimate the host stellar mass.

@@

Considering the simulations by~\citet{Ding2017a} and the fact that we are able to obtain very consistent host magnitudes for the HE0435, RXJ1131 and WFI2033 using the independent approaches, the fidelity of the inferred magnitude is expected to be well characterized by the quoted uncertainties.
\blue{
Our systems do not have sufficient multi-band information to constrain their metallicity and star formation histories. Therefore, we adopted a simple stellar population template to calculate \mstar, but reflect the lack of information in the associated uncertainty. Following \citet{Bell2001}, we estimate a typical uncertainty of 0.2~dex for all the \mstar\ measurements in this paper.}

@@

\blue{
Note that this apparent evolution is obtained directly using the observed sample, before considering selection effects~\citep{Tre++07,Schulze2011,Bennert++2011, Schulze2014,Park15}. For instance, adopting the framework introduced in~\citet{Schulze2015}, D20 estimated that accounting for selection biases would yield a more modest evolution $\gamma=0.50\pm0.25$. The lensed AGN systems considered in this work were selected for time-delay cosmography, based on the availability of a time delay and the known detectability of the host galaxy. This is a complex selection function, different than the one used by D20. Although it is encouraging that the two samples present the same apparent evolution, inferring the true underlying evolution requires modelling the selection function, as studied by D20. A full modelling of the complex selection function is not warranted given the small size of the lensed quasars sample. Thus, at this stage, the lensed quasars should be considered as a check on possible systematic measurement error in the D20 analysis rather than a stand-alone measurement.}

@@

We directly compare our sample to the recent measurements by~\citet[][D20]{Ding2020a}, who used the same approach to derive the \mstar\ and calibrated the \mbh\ using consistent recipes. The \mbh-\mstar\ correlation and its evolution with cosmic time {\bf are} in excellent agreement with the results obtained by D20, as shown in Figure~\ref{fig:scaling_relation}. \blue{Currently, the sample of lensed AGN (i.e., eight systems) is quite limited in size; as a result, given the statistical errors it is not worth modelling selection effect.} Nevertheless, the good agreement with the D20 sample strengthens their conclusions, which can be summarized as follows. First, the growth of the supermassive black hole \blue{somewhat} predates that of its host galaxy during their co-evolution, even when considering total stellar mass, as it is often done at high redshift. However, the actual morphologies of these hosts are likely to be more complex, including a disk and bulge component, as in the case of  RXJ1131. In contrast, the galaxies in the local sample are typically bulge-dominated and the bulge mass is adopted as their  \mstar. Thus, the reported evolution is weaker than what would be inferred by comparing \mbh\ to bulge \mstar\ at all redshifts \citep{Bennert++2011}. Taken together, these results are consistent with a scenario in which the stellar mass is transferred from disk to the bulge at a faster rate than the growth of \mbh\ since $z\sim2$.
